Playing Games: A Case Study in Active Learning Applied to Game Theory

### 2015-06-29 paper/paper.tex
Updating document and plot
- Including a plot showing results only for students who reduce their
  guess
- Adding to text describing this.

\usepackage{fullpage} % Package to use full page
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}

@@

\begin{itemize}
    \item Section \ref{sec:best_responses} will describe an in class activity
        and software package used to introduce students to the topic of best
        response dynamic \cite{Maschler2013}.
    \item Section \ref{sec:repeated_games} will describe an implementation of
        Axelrod's tournament \cite{Axelrod1980a,Axelrod1980a}.
\end{itemize}

@@

To make use of this game in class as an introduction to the concept of best
response dynamics students are handed a sheet of paper inviting them to write
down a first play. After this initial play, a discussion is had that
demonstrates that the equilibrium for this game is for all players to guess 0.
This is shown diagrammatically in Figure
\ref{fig:rationalisation_of_two_thirds_game}.
Following this discussion students are invited to play again and write down
their second guess.

@@

>>>>

\usepackage{fullpage} % Package to use full page
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{subcaption}

@@

\begin{itemize}
    \item Section \ref{sec:best_responses} will describe an in class activity
        and software package used to introduce students to the topic of best
        response dynamic \cite{Maschler2013}.
    \item Section \ref{sec:repeated_games} will describe an implementation of
        Axelrod's tournament \cite{Axelrod1980a, Axelrod1980b}.
\end{itemize}

@@

To make use of this game in class as an introduction to the concept of best
response dynamics students are handed a sheet of paper inviting them to write
down a first play. After this initial play, a discussion is had that
demonstrates that the equilibrium for this game is for all players to guess 0.
This is shown diagrammatically in Figure
\ref{fig:rationalisation_of_two_thirds_game}.

Following this discussion students are invited to play again and write down
their second guess. All of the results are collected, the author has used paper
forms but an automated approach could also be used. In general the input and
analysis of the data takes less than 10 minutes and can be done by a helper
during another class activity.
Following this, the result shown in Figure \ref{fig:histogram_of_guess} are
shown and discussed with the class.

@@

The author has used this activity on a large number of occasions and at all
times collected the data. Figure \ref{fig:histogram_of_guess} shows the
distribution of the guesses (depending on the round of play):

\begin{figure}[htpb]
    \begin{subfigure}{.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/histogram_of_guesses.pdf}
        \caption{Frequency of guesses depending on the round of play.}
        \label{fig:histogram_of_guess}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/jointplot_of_guesses.pdf}
        \caption{Linear relationship between guesses of each round of play.}
        \label{fig:jointplot_of_guess}
    \end{subfigure}
    \caption{Results from all data collected.}
    \label{fig:all_results}
\end{figure}

We see that the second round (after the rationalisation of play described in
Figure \ref{fig:rationalisation_of_two_thirds_game}) has guesses that are closer
to the expected equilibrium behaviour.
Figure \ref{fig:jointplot_of_guess} confirms this showing the linear
relationship (albeit a weak one with \(R^2=.2\)):

\begin{equation}
    (\text{Second guess}) = .203\times(\text{First guess}) + 9.45
    \label{eq:linear_relationship}
\end{equation}

The fact that the coefficient of the relationship is less than one highlights
that the second guess is in general lower than the first guess. As can be seen
in Figure \ref{fig:all_results} not all students reduce their guess. Figure
\ref{fig:results_with_decreasing_guess} shows the results when removing these
irrational moves. In this particular case the linear relationship is in fact
stronger \(R^2=.43\):

\begin{equation}
    (\text{Second guess}) = .33\times(\text{First guess}) + 0.20
    \label{eq:linear_relationship}
\end{equation}

\begin{figure}[htpb]
    \begin{subfigure}{.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/histogram_of_guesses_removing_increasing_guesses.pdf}
        \caption{Frequency of guesses depending on the round of play.}
        \label{fig:histogram_of_guess}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/jointplot_of_guesses_removing_increasing_guesses.pdf}
        \caption{Linear relationship between guesses of each round of play.}
        \label{fig:jointplot_of_guess}
    \end{subfigure}
    \caption{Results from data when removing increasing guesses.}
    \label{fig:results_with_decreasing_guess}
\end{figure}

At the end of the activity, students are shown the graphical results and a
discussion had about why the theoretic equilibrium was not the winner. This
discussion usually revolves around the observation that not everyone acted
rationally and second that some participants felt like they should `spoil' the
game by guessing larger in the second round.

Finally, if time permits (and depending on the level of the participants), the
linear relationship of (\ref{eq:linear_relationship}) is used to discuss what
would happen if more rounds were to be played. In particular it is possible to
discuss ideas of convergence when generalising (\ref{eq:linear_relationship}) to
be:

\begin{equation}
    \text{Guess}_{n+1} = .203\times\text{Guess}_n + 9.45
    \label{eq:extrapolated_linear_relationship}
\end{equation}
